---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rhealthdiag

<!-- badges: start -->
<!-- badges: end -->

*The goal of rhealthdiag is a package developed in the department of prevention and control of diseases transmitted by vector of the [Secretary of Health of Veracruz](https://www.ssaver.gob.mx/) and with colaboration of the federal level.*

### **overview**

**rhealthdiag** is a package designed to generate a health diagnosis of the mortality of a state, municipality or locality using data from the [INEGI](https://www.inegi.org.mx/programas/mortalidad/#Datos_abiertos). The package contain the mortality of Puebla state (with the updates we add all states) and the functions for generate the tables, static and interactive maps.


**Datasets**

  - **`mortalidad_general_21`** Mortality of Puebla State (2012_2019).
  - **`mortalidad_general_12`** Mortality of Guerrero State (2012_2019).

**Functions**

  - **`mort_year_edo_dt`** Generates tables of causes of death from 2012 to 2019 at the state level.
  
  - **`mort_year_mpo_dt`** generates tables of causes of death from 2012 to 2019 at the municipality level.
  
  - **`mort_year_loc_dt`** Generates tables of causes of death from 2012 to 2019 at the locality level.

## Installation

You can install the released version of healthdiagr from [CRAN](https://CRAN.R-project.org) with:

``` r
#install.packages("rhealthdiag")
```

## Development version

To get a bug fix, or use a feature from the development version, you can install boldenr from GitHub.

``` r
# install.packages("remotes")
remotes::install_github("fdzul/rhealthdiag")
```
or

``` r
# install.packages("devtools")
devtools::install_github("fdzul/rhealthdiag")
```

## Example

This is a basic example which shows you how load the mortality dataset in R:

```{r example}
library(rhealthdiag)
head(rhealthdiag::mortalidad_general_21)
```

## Authors

* **Felipe Antonio Dzul Manzanilla** -**https://github.com/fdzul** - Packages developed in github:

  1) [denhotspots](https://github.com/fdzul/denhotspots). 
  2) [boldenr](https://github.com/fdzul/boldenr). 
  3) [dendata](https://github.com/fdzul/dendata).
  4) [deneggs](https://github.com/fdzul/deneggs).
  5) [rgeomex](https://github.com/fdzul/rgeomex).
  6) [covid19mx](https://github.com/fdzul/covid19mx).


See also the list of [contributors](https://github.com/fdzul/deneggs/contributors) who participated in this project.

## License

This project is licensed under the MIT License - see the [LICENSE.md](LICENSE.md) file for details


## Inspiration

The package was inspired by the need to contribute to generate health diagnoses at the state, municipality and local level.

## Getting help

If you encounter a clear bug, please file a minimal reproducible example
on [github](https://github.com/fdzul/healthdiagr/issues). For questions
and other discussion, please feel free to contact me (felipe.dzul.m@gmail.com)

-----

Please note that this project is released with a [Contributor Code of
Conduct](https://dplyr.tidyverse.org/CODE_OF_CONDUCT). By participating
in this project you agree to abide by its terms.

